---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
```

## INLAloggrowth

This is a R package to implement the Spatial Logistic Growth model described in
[paper]. It uses the `cgeneric` interface in `R-INLA` to implement the model in `C` and uses `inlabru` to fit the spatial models.

## Installation
The 'INLA' package is needed for fitting a model. You can install it with
```{r eval = !require(INLA)}
install.packages("INLA",repos=c(getOption("repos"),INLA="https://inla.r-inla-download.org/R/testing"), dep=TRUE) 
```

You can install the latest version of INLAloggrowth from [GitHub](https://github.com/ljb34/INLAloggrowth) with
```{r ghinst, eval = FALSE}
## install.packages("remotes")
remotes::install_github("ljb34/INLALogisticGrowth")
```

## A simple example

Loading packages
```{r library}
library(INLA)
library(inlabru)
library(INLAloggrowth)
library(sf)
library(fmesher)
```

Simulate logistic growth data
```{r simulate, results = FALSE}
simdata <- simulate_loggrowth(growth = 1, carry.cap = 500, movement = 0.15, sigma = 0.35,
                               initial.pop = 150, initial.range = 0.15, initial.sigma=0.075,
                               timesteps = 3,sample.type = "LGCP", boundaries = c(0,1), debug = T,
                               max.edge = 0.25)
library(ggplot2)
#Plot observations
ggplot()+
  gg(simdata$animal_obs)+
  facet_wrap(~time)+
  ggtitle("Observed Points")

#Plot unobserved field
ggplot()+
  gg(simdata$animal, aes(fill = exp(field)), geom = "tile")+
  scale_fill_viridis_c()+
  facet_wrap(~time)+
  ggtitle("Abundance")
```



Define spatial and temporal discretization meshes
```{r meshes}
bnd <- spoly(data.frame(easting = c(0,1,1,0), northing = c(0,0,1,1)), format = "sf")
mesh_space <- fm_mesh_2d(boundary = bnd,
                       max.edge = 0.25)
mesh_time <- fm_mesh_1d(
  loc = 0:3)
```

Fit a simple model to the first year. The output of this model becomes the prior for the initial condition. 

```{r first-model}
matern <- inla.spde2.pcmatern(mesh_space,
                              prior.sigma = c(0.5, 0.01),
                              prior.range = c(0.1, 0.02))
cmp <- geometry ~ smooth(geometry, model = matern) +
  initial(1,model = "linear")-1

first_fit <- bru(cmp, simdata$animal_obs[simdata$animal_obs$time == 1,],
                 family = "cp", domain = list(geometry = fm_subdivide(mesh_space,1)))
```

Extract mean and precision matrix

```{r extract}
index <- min(which(stringr::str_sub(rownames(first_fit$summary.fitted.values),8,8)!= "A"))
n.nodes <- first_fit$misc$configs$nconfig
nodes <- data.frame(log.prob=rep(NA,n.nodes))
mat_list <- list()
mean_list <- list()
for(i in 1:n.nodes){
  nodes[i,]<- first_fit$misc$configs$config[[i]]$log.posterior
  Q <- first_fit$misc$configs$config[[i]]$Q[1:mesh_space$n, 1:mesh_space$n]
  dQ <- diag(Q)
  Q <- Q + Matrix::t(Q)
  diag(Q) <- dQ
  mat_list[[i]] <- Q
  mean_list[[i]] <- first_fit$misc$configs$config[[i]]$improved.mean[1:mesh_space$n]
}
nodes <- dplyr::mutate(nodes, weight = exp(log.prob)) %>%
  dplyr::mutate(weight.prob = weight/sum(weight))
Q <- Reduce("+", Map(function(m,w) w*m, mat_list,nodes$weight.prob))
initial.precision <-Diagonal(mesh_space$n,
                             1/((first_fit$summary.fixed$sd**2)+(first_fit$summary.fitted.values$sd[index-1 +1:mesh_space$n]**2)))%*%Q
weighted.means <- Map(function(v,p) v*p, mean_list, nodes$weight.prob)
prior.mean <- Reduce("+", weighted.means) + first_fit$summary.fixed$mean[1]
```

Fit logistic growth model

```{r Logistic, results = FALSE}
priors = list(cc = c(log(500), 0.5), growth = c(log(1), 0.5), 
              movement = c(log(0.15),0.5), sigma = c(log(0.5),0.5))
logistic_fit <- iterate.fit(formula = geometry + time ~ -1, #Need to remove intercept
                         data = simdata$animal_obs, family = "cp",
                         domain = list(geometry = fm_subdivide(mesh_space,1), time = 1:3),
                         smesh = mesh_space, tmesh = mesh_time, samplers = bnd,
                         prior.mean = prior.mean, 
                         prior.precision = Q,
                         priors = priors, initial.growth=log(1),
                         initial.carry.cap=log(500), initial.move.const = log(0.15),
                         initial.sigma = log(0.5), gamma = 0.5,
                         options = list(verbose = F, control.inla = list(strategy = "gaussian", int.strategy = "eb")))
```
Parameters on log scale. In order growth, carrying capacity, dispersal, sigma
```{r output}
summary(logistic_fit$fit) 
```
Parameters on actual scale. In order growth, carrying capacity, dispersal, sigma
```{r Parameters}
exp(logistic_fit$fit$summary.hyperpar) 
```